GCD_Assignment
==============

## Setting the work directory, importing datasets and opening information files

```{r,echo=TRUE}
# The following code chunk is used to set the working directory of R in a
# way that the system can import the datasets referent to the train and
# test values and, with them, import the files related to the labels
# of activities and variables

setwd("UCI HAR Dataset")
xVector <- read.table("features.txt")
xActivity <- read.table("activity_labels.txt")
setwd("test")
xTest <- read.table("X_test.txt")
xTst_label <- read.table("y_test.txt")
xTst_subject <- read.table("subject_test.txt")
setwd("../")
setwd("train")
xTrain <- read.table("X_train.txt")
xTrn_label <- read.table("y_train.txt")
xTrn_subject <- read.table("subject_train.txt")
setwd("../")
```

## Cleaning the names of the labels for the "special" vector

```{r,echo=TRUE}
# On the chunk below, I used regex to remove from the variable names some elements
# that could make them less readable, in a way that the data coulde be more easily
# understood
xVector$V2 <- gsub("\\(|\\)|\\-",",",xVector$V2)
xVector$V2 <- gsub("\\,", ".", xVector$V2)
```

## Setting the name of the columns related to the "special" vector

```{r,echo=TRUE}
colnames(xTest) <- c(xVector$V2)
colnames(xTrain) <- c(xVector$V2)
```

## Subsetting the datasets with only the colums where the elements were related to the mean or standard deviation of the values colected

```{r,echo=TRUE}
# On the lines below I generated a vector containing the indexes of all the variables
# of the "special" vector that were related to mean and standard deviation and used
# this vector to subset the original datasets in new ones containing only those that
# are present on the vector
usableName_vector <- grep("mean|std", xVector$V2)
xTest <- xTest[,usableName_vector]
xTrain <- xTrain[,usableName_vector]
```

## Setting the name of the columns related to the labels of activities and to the the subjects that participated in the studies

```{r,echo=TRUE}
colnames(xTrn_label) <- "label"
colnames(xTrn_subject) <- "subject"
colnames(xTst_label) <- "label"
colnames(xTst_subject) <- "subject"
```

## Binding columns and merging datasets

```{r,echo=TRUE}
# in this chunk I generated 2 columns in every dataset with the label for the activity
# in each vector and the code for which subject generated that data. Then, I merge
# the datasets by binding the rows of the train and test datasets
xTest <- cbind(xTest,xTst_subject)
xTest <- cbind(xTest,xTst_label)
xTrain <- cbind(xTrain,xTrn_subject)
xTrain <- cbind(xTrain,xTrn_label)
xData <- rbind(xTrain,xTest)
```

## Generating a new line and naming it's elements by factor

```{r,echo=TRUE}
xData$activity <- xData$label
xData$activity <- factor(xData$activity, labels=xActivity$V2)
```

## Generating the new tidydata file

```{r,echo=TRUE}
# in the lines below I generate a new dataset with the tidydata related to the average
# value for each variable of the "special" vector for every activity realized by
# every subject. This is done by agreggating the values by subject, label and activity.
# After that, I export the tidydata to a .txt file.
tidyData<-aggregate(. ~ subject + label + activity, data = xData, mean)
setwd("../")
write.table(tidyData, "tidyData.txt", sep=" ")